Exploring One Variable with R
========================================================

### Load the Pseudo-Facebook User Data

```{r Pseudo-Facebook User Data}
pf <- read.csv('pseudo_facebook.tsv',sep = '\t')
names(pf)
```

***

### Histogram of Users' Birthdays

```{r Histogram of Users\' Birthdays}
library(ggplot2)

names(pf)
qplot(x = dob_day, data = pf)

qplot(x = dob_day, data = pf) +
  scale_x_discrete(breaks=1:31)
```

***
### Faceting
Notes: This takes a ~ and a variable.  Here we are plotting one histogram for each month in the same plot.

```{r Faceting}
qplot(x = dob_day, data = pf) + 
  scale_x_discrete(breaks = 1:31) +
  facet_wrap(~dob_month, ncol = 3)
```

Notes: This is similar to facet_grid(). This takes as a parameter a formula in the format: (vertical variable ~ horizontal variable).

See the [facet documentation](http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/).

***

#### Create a histogram of friend counts
Basic Plot
```{r Friend Count}
qplot(x = friend_count, data = pf)
```
This limits the x axis to 1000 so we can see more detail in the bulk of the data.  2 different methods for the same plot.
```{r Friend Count limited x}
qplot(x = friend_count, data = pf, xlim = c(0, 1000))
qplot(x = friend_count, data = pf) +
  scale_x_continuous(limits = c(0, 1000))
```

### Faceting Friend Count
Create a histogram by gender breaking up the x axis by 50 units
```{r Faceting Friend Count 1}
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender, ncol=2)
```
#### What?  Why are there 3 histograms?
#### Let's subset the data 2 different ways
```{r Faceting Friend Count 2}
# This removes all data with gender as NA
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender, ncol=2)

# This removes all NA values for any field
qplot(x = friend_count, data = na.omit(pf), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender, ncol=2)
```
#### Ok, that is better.

***
### Statistics 'by' Gender

```{r Statistics \'by\' Gender}
table(pf$gender)

# the by function takes a variable, a categoricical variable and a function as parameters
by(pf$friend_count, pf$gender, summary)
```

On average women have more friends.
In a long tailed distribution like this one, the median is a better measure of central tendency than the mean.
***

### Tenure
Notes:

```{r Tenure}
# Using qplot
qplot(x=tenure/365, data=pf, 
      xlab='Number of years using Facebook',
      ylab = 'Number of users in sample',
      binwidth=0.5,
      color=I('black'), fill=I('#099D09')) + 
    scale_x_continuous(breaks=seq(1,7,1), limits=c(0,7))
```

***

#### Create a histogram of tenure by year?

```{r Tenure Histogram by Year}
# Using ggplot
ggplot(aes(x = tenure / 365), data = pf) + 
  geom_histogram(color = 'black', fill = '#F79420') + 
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7)) +
  xlab('Number of years using Facebook') + 
  ylab('Number of users in sample')
```

***

### User Ages

```{r User Ages}
qplot(x=age, data=pf, binwidth = 1,
      fill=I('#5760AB')) +
  scale_x_discrete(breaks = seq(0,113,5))
      
```
***

### Transforming Data
Plot friend counts
```{r}
qplot(x = friend_count, data = pf)
```
#### This has a long right tail.  Let's try the log of friend_count.
```{r}
qplot(x = log10(friend_count), data = pf)
summary(log10(pf$friend_count))
```
#### So a mean of -Inf is weird, right?
#### Let's add 1 to all of the friend counts to fix this.
```{r}
summary(log10(pf$friend_count + 1))
```
#### We could also see the square root of friend count.
```{r}
summary(sqrt(pf$friend_count))
```

#### Create 3 plots for friend count, log10 of friend count + 1 and sqrt of friend count
```{r}
library(gridExtra)
p1 = qplot(x=friend_count, data=pf)
p2 = qplot(x=log10(friend_count+1), data=pf)
p3 = qplot(x=sqrt(friend_count), data=pf)

grid.arrange(p1,p2,p3, ncol=1)
```
#### The same plots using ggplot
```{r}
p1 = ggplot(aes(x=friend_count), data=pf) + 
  geom_histogram()
p2 = p1 + scale_x_log10()
p3 = p1 + scale_x_sqrt()
grid.arrange(p1, p2, p3, ncol=1)
```

***

### Add a Scaling Layer
This plots the same plot using qplot and ggplot.  Notice that the x axis lable is different.  In ggplot, you need to set the x label to accurately indicate the scale.

```{r Add a Scaling Layer}
logScale = qplot(x=log10(friend_count), data=pf)
countScale = ggplot(aes(x=friend_count), data=pf) +
  geom_histogram() +
  scale_x_log10()
grid.arrange(logScale, countScale, ncol=2)
```

***

### Frequency Polygon
We can plot likes by gender using a frequency polygons
```{r Likes on the Web}
qplot(x=www_likes, data=subset(pf,!is.na(gender)),
      geom='freqpoly', color=gender) +
  scale_x_continuous() #+
  #scale_x_log10()

ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) + 
  scale_x_log10()

by(pf$www_likes, pf$gender, sum)
```
### We can also change the y axis to a proportion
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), 
      binwidth = 10) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50)) +
  facet_wrap(~gender)

qplot(x = friend_count, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)),
      xlab = 'Friend Count',
      ylab = 'Proportion of Users with that friend count',
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50))
```
#### Now let's see where women overtake men in likes
```{r}
qplot(x = friend_count, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)),
      xlab = 'Friend Count',
      ylab = 'Proportion of Users with that friend count',
      binwidth = 10, geom = 'freqpoly', color = gender) +
  scale_x_continuous(lim = c(800, 1000), breaks = seq(0, 1000, 50))

by(pf$www_likes, pf$gender, sum)
```

***

### Box Plots
Always pass the categorical variable for x and the continuous variable for y.
```{r Box Plots}
qplot(x=gender, y=friend_count, 
      data=subset(pf, !is.na(gender)),
      geom='boxplot') +
  coord_cartesian(ylim=c(0,250))
```
#### Use coord_cartesian to limit the y data visually, but keep the data points in the analysis.
***

### Box Plots, Quartiles, and Friendships

```{r Box Plots, Quartiles, and Friendships}
qplot(x=gender, y=likes_received, 
      data=subset(pf, !is.na(gender)),
      geom='boxplot') +
  coord_cartesian(ylim = c(0, 200))

by(pf$friend_count, pf$gender, summary)
by(pf$friendships_initiated, pf$gender, summary)
```

***
### Getting Logical

```{r Getting Logical}
summary(pf$mobile_likes)
summary(pf$mobile_likes > 0)

mobile_check_in = NA
pf$mobile_check_in = ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in = factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```